```{r models}
# code written by Paul Tanger, Brook Moyers or John Lovell. Please cite appropriately.

setwd("../input")
#############################
# list approach - store all models in a list and use lapply to do things to all

load("DS2013AllMeansNewWithCTDFixHI.Robject")
DS2013AllMeans$phenotype = as.factor(DS2013AllMeans$phenotype)
DS2013AllMeans$phenotype  <- factor(DS2013AllMeans$phenotype , levels = 
                                       c("DTF",
                                         "DTH",
                                         "biomass",
                                         "height",
                                         "grain",
                                         "HI",
                                         "CTD",
                                         "Chla",
                                         "HTPheight",
                                         "NDRE",
                                         "NDVI",
                                         "NIR_760",
                                         "R_670",
                                         "RE_730"))
DS2013AllMeans$pheno_rep_date = paste(DS2013AllMeans$phenotype, DS2013AllMeans$rep, DS2013AllMeans$date, sep="_")

# get rid of raw phenos and dates we don't want
rawspectra = c("NIR_760", "R_670", "RE_730", "grain", "HI")
DS2013AllMeans2 = DS2013AllMeans
DS2013AllMeans2 = subset(DS2013AllMeans2, !(phenotype %in% rawspectra))

DS2013AllMeans2 = droplevels(DS2013AllMeans2)

# save it for modeling
as.data.frame(colnames(DS2013AllMeans2))
DS2013AllMeans3 = DS2013AllMeans2[,c(1,2,3,4,8,5)]
alldata = DS2013AllMeans3

# make it wide.. by rep..
alldatabyrepdate = dcast(alldata[,c(1,2,3,4,6)], line + rep + date ~ phenotype , value.var="mean", mean, na.rm=T)

alldata = alldatabyrepdate

# make list of phenotypes
varlist = names(alldata)[4:ncol(alldata)]

# make line and rep a factor
alldata$line = as.factor(alldata$line)
alldata$rep = as.factor(alldata$rep)

# rename to make rest of code compatible..
colnames(alldata)[colnames(alldata)=="line"] = "Line"

# run random line and rep models
models = lapply(varlist, function(x) { 
  lmer(substitute(i ~ + (1 | Line) + (1 | rep), list(i=as.name(x))), data=alldata)
})

# name models in list..
names(models) = varlist

###################################################################################
# examine models
###################################################################################

lapply(models, summary)
lapply(models, AIC)
lapply(models, getAICc)
lapply(models, BIC)

###################################################################################
# get stuff from models
###################################################################################

varcomps = lapply(models, varcompsdf)
varcompsall = ldply(varcomps, data.frame)
# make new column for excel
varcompsall$H2 = varcompsall$varPCT / 100
# rename col
colnames(varcompsall)[colnames(varcompsall)==".id"] = "phenotype"
# round values
varcompsall$H2 = formatC(round(varcompsall$H2, digits=2), 2, format="f")

# export it
setwd("../output")
filename = addStampToFilename("DS2013VarComps", "tsv")
write.table(varcompsall, file=filename, sep="\t", quote=F, row.names=F)

```

```{r fixedmodels, eval=F}
###################################################################################
# model phenotypes with line and rep fixed ... to get LSmeans
###################################################################################
# disable scientific notation
options(scipen=999)

alldata$Line = as.factor(alldata$Line)

# get fixed effect models.. line and rep fixed..
fixedmodels = lapply(varlist, function(x) { 
  lm(substitute(i ~ Line + rep, list(i=as.name(x))), data=alldata)
})

# name models in list
names(fixedmodels) = varlist

###################################
# updated for lsmeans version 2
# see ref.grid class definition here around page 22
# http://cran.r-project.org/web/packages/lsmeans/lsmeans.pdf

# get lsmeans
fixedmodels.lsmeans = lapply(fixedmodels, function(x) { lsmeans(x, ~ Line)})
# convert list to dfs
listlength = length(fixedmodels.lsmeans)

# define empty df
lsmeansdf = as.data.frame(setNames(replicate(7,numeric(0), simplify = F), c("Line", "lsmeans", "SE", "df", "lowerCI", "upperCI", "resp")))
i=1
# this error is ok.. just ignore it...
#Error in make.link(misc$tran) : ‘+’ link not recognised
for (i in 1:listlength) {
   lsmeansdf = rbind(lsmeansdf, cbind(summary(fixedmodels.lsmeans[[i]]), list(resp=getrespname(fixedmodels[[i]]))))
}

# dcast it
lsmeanswide = dcast(lsmeansdf, Line ~ resp, value.var="lsmean")
# rename cols
varlistlsmeans = lapply(varlist, function(x) {paste(x, "_LSmeans", sep="")})
names(lsmeanswide) = c("Line", varlistlsmeans)
colnames(lsmeanswide)[colnames(lsmeanswide)=="Line"] = "ID"
filename = addStampToFilename("DS2013FinalLSmeanswithparents", "csv")
#write.table(lsmeanswide, file=filename, sep=",", quote=F, row.names=F)